Inside Out: Externalizing Assumptions in Data Analysis as Validation Checks

### 2024-12-12 index.qmd
fix a reference

The typical approach to studying the association between PM10 and mortality is to apply a generalized linear model with a Poisson link to relate daily mortality counts to daily measures of PM10. Based on previous work and the range of effect sizes published in the literature, an analyst might expect the coefficient for PM10 in this GLM to lie between $[0, 0.005]$, after adjusting for daily temperature [@samet200fine; @welty2005acute]. Observing an estimated coefficient outside of this interval would be highly unusual and would warrant a serious re-examination of the analysis process. Therefore, our unexpected outcome in this analysis will be the binary indicator of whether the estimated coefficient for PM10 in the GLM lies outside of the interval $[0, 0.005]$. In addition to providing a more substantial problem for our methods, this example also demonstrates how the procedure presented in @sec-method can be used to select cutoff values in the analysis checks to diagnose an unexpected PM10 coefficient from the generalized linear model.

>>>>

The typical approach to studying the association between PM10 and mortality is to apply a generalized linear model with a Poisson link to relate daily mortality counts to daily measures of PM10. Based on previous work and the range of effect sizes published in the literature, an analyst might expect the coefficient for PM10 in this GLM to lie between $[0, 0.005]$, after adjusting for daily temperature [@samet2000fine; @welty2005acute]. Observing an estimated coefficient outside of this interval would be highly unusual and would warrant a serious re-examination of the analysis process. Therefore, our unexpected outcome in this analysis will be the binary indicator of whether the estimated coefficient for PM10 in the GLM lies outside of the interval $[0, 0.005]$. In addition to providing a more substantial problem for our methods, this example also demonstrates how the procedure presented in @sec-method can be used to select cutoff values in the analysis checks to diagnose an unexpected PM10 coefficient from the generalized linear model.
